The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation

### 2023-02-09 paper.tex
prior art portability
message supersed paragraph

In \yaq{}, communication between daemons and clients is performed over TCP/IP using Apache Avro RPC \cite{AvroSpecification}.
Avro provides an agreed upon standard for efficient serialization of data and method calls from a remote (client) process.
Practically, the \yaq{} interface looks like a collection of methods or functions, which Avro calls ``messages''.
Each message has defined input parameters and output return types.
A sensor might implement a message called ``\texttt{get\_measured}'' which takes no parameters and returns a dictionary mapping channel names to numeric or array data.
A motor would implement a pair of messages for setting and reading back the motor position: ``\texttt{set\_position(float position)} $\rightarrow$ \texttt{null}'' and ``\texttt{get\_position()} $\rightarrow$ \texttt{float}''.
Messages make up the lowest level functionality of the \yaq{} interface.
Each individual communication between client and daemon involves one message being requested by the client and the response returned from the daemon.
If multiple messages are sent simultaneously, the daemon processes one message at a time.
The protocol is preemptive: in the case of a conflict, the daemon will obey the last message to arrive.
Each daemon supports a collection of messages for its unique functionality, called the ``protocol''.
When a client first connects to a daemon over TCP/IP, the daemon provides a complete description of its own protocol.
We therefore refer to the daemons as ``self-describing''.

@@

In a \yaq{} context, new hardware can be incorporated into an instrument through the addition of a new daemon.
The \yaq{} architecture simplifies hardware interface development in several ways.
First, because daemons are separate and portable programs, the development effort can be spread across the community of \yaq{} users.
Often, researchers can download an existing daemon rather than writing a new one.
Second, \yaq{} daemon development can be performed separately from the particulars of any individual client.
Often, this separation allows initial hardware enablement work to be done on a researchers personal machine before the new hardware peripheral is installed in the instrument.
Third, when developing trait-compliant protocols, it becomes easy to design and fully test your hardware interface.
Traits are unambiguous and well-described, making an obvious target for development.
Tooling exists to verify full trait compliance, for example you can use \texttt{yaqc-qtpy} to provide a graphical program to interact with your hardware immediately.
Finally, as discussed in Section II, \yaq{} provides options to design using remote hardware or unusual interfaces when necessary.

>>>>

In \yaq{}, communication between daemons and clients is performed over TCP/IP using Apache Avro RPC \cite{AvroSpecification}.
Avro provides an agreed upon standard for efficient serialization of data and method calls from a remote (client) process.
Practically, the \yaq{} interface looks like a collection of methods or functions, which Avro calls ``messages''.
Each message has defined input parameters and output return types.
A sensor might implement a message called ``\texttt{get\_measured}'' which takes no parameters and returns a dictionary mapping channel names to numeric or array data.
A motor would implement a pair of messages for setting and reading back the motor position: ``\texttt{set\_position(float position)} $\rightarrow$ \texttt{null}'' and ``\texttt{get\_position()} $\rightarrow$ \texttt{float}''.
Messages make up the lowest level functionality of the \yaq{} interface.
Each daemon supports a collection of messages for its unique functionality, called the ``protocol''.
When a client first connects to a daemon over TCP/IP, the daemon provides a complete description of its own protocol.
We therefore refer to the daemons as ``self-describing''.

Each individual communication between client and daemon involves one message being requested by the client and the response returned from the daemon.
If multiple messages are sent simultaneously, the daemon processes one message at a time.
The protocol is preemptive: in the case of a conflict, the daemon will obey the last message to arrive.
By being completely open to messages from multiple clients, \yaq{} makes it possible to accidentally supersede a message.
This has not become a problem for our usage where instruments are controlled by single scientists who can clearly see all of the clients interacting with their system.
Typically there is just one client sending `set\_position` and many clients polling `get\_position` to record data or display real-time information.
Users may choose to implement access control systems, but we do not foresee building access control into the \yaq{} protocol.
This is similar to how access control works in practice at some large facilities.

@@

In a \yaq{} context, new hardware can be incorporated into an instrument through the addition of a new daemon.
The \yaq{} architecture simplifies hardware interface development in several ways.
First, because daemons are separate and portable programs, the development effort can be spread across the community of \yaq{} users.
Often, researchers can download an existing daemon rather than writing a new one.
Second, \yaq{} daemon development can be performed separately from the particulars of any individual client.
Often, this separation allows initial hardware enablement work to be done on a researchers personal machine before the new hardware peripheral is installed in the instrument.
Separability, portability, and distributed development are advantages common to many open-source hardware interface projects.
Third, when developing trait-compliant protocols, it becomes easy to design and fully test your hardware interface.
Traits are unambiguous and well-described, making an obvious target for development.
Tooling exists to verify full trait compliance, for example you can use \texttt{yaqc-qtpy} to provide a graphical program to interact with your hardware immediately.
Finally, as discussed in Section II, \yaq{} provides options to design using remote hardware or unusual interfaces when necessary.
